---
title: "Data612 P3"
author: "Eunkyu Hahm"
date: "6/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project 3 | Matrix Factorization methods

To implement a matrix factorization method—such as singular value decomposition (SVD) or Alternating Least Squares (ALS)—in the context of a recommender system.

### 1. Load libraries
```{r}
library(recommenderlab)
library(reshape2)
library(knitr)
library(tidyverse)
```


### 2. Load data and subset data
```{r}
data(MovieLense)
MovieLense
```

Subsetting users have rated at least 50 movies and movies have been watched more than 100 times.
```{r}
ratings_m <- MovieLense[rowCounts(MovieLense)>50, colCounts(MovieLense)>100]
ratings_m
```

### 3. train and test datasets

```{r}
#1 is users in train dataset, 0 is in test datset#
which_train<- sample(x = c(1, 0), size = nrow(ratings_m),replace = TRUE, prob = c(0.8, 0.2))
head(which_train)
```
Train dataset   
Test dataset
```{r}
train_m <- ratings_m[which_train,]
test_m <- ratings_m[!which_train,]
train_m
test_m
```

### 4. Singular Value Decomposition
```{r}
svd_model <- Recommender(data=train_m, method="SVD", parameter= list(k=20))
svd_model
```

```{r}
svd_predict <- predict(object = svd_model, newdata = test_m, n=6)
svd_predict
```
Recommendation for the first five user. 
```{r}
svd_predict@items[1:5]
```

