---
title: "Data6012 project 1"
author: "Eunkyu Hahm"
date: "6/7/2020"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: TRUE
      smooth_scroll: False
    number_sections: true
    theme: united
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Project 1

## Hotel recommendation system
This system recommends the chain hotesl to travelers. 

## Dataset
```{r}
#Using toy dataset
hotels <- matrix(c(4,4,4,3,5,4,4,3,3,5,2,3,3,4,2,4,3,4,5,4,5,5,2,5,4,NA,3,4,3,4,4,5,NA,2,2,3), ncol = 6, byrow = TRUE)
colnames(hotels)<- c("Marriott","Hilton","Best Western", "Hyatt", "Intercontinental","Wyndham")
rownames(hotels)<- c("user1","user2","user3","user4","user5","user6")
hotels
```

## Separating training and test datasets
```{r}
training <- matrix(c(4,NA,4,3,5,4,4,3,NA,5,2,3,3,4,2,4,NA,4,NA,4,5,5,2,5,4,NA,3,NA,3,4,4,5,NA,2,2,NA),ncol = 6, byrow = TRUE)
colnames(training)<-c("Marriott","Hilton","Best Western", "Hyatt", "Intercontinental","Wyndham")
rownames(training)<- c("user1","user2","user3","user4","user5","user6")

test <- c(5,4,3,4,3,3)
```

## Raw average (mean) rating
```{r}
raw_average <- mean(training, na.rm=TRUE)
raw_average
```

## Calculating the RMSE

RMSE - training dataset
```{r}
rmse_training <- sqrt(mean((training-raw_average)^2, na.rm= TRUE))
rmse_training
```

RMSE - test dataset
```{r}
rmse_test <- sqrt(mean((test-raw_average)^2))
rmse_test
```

## Calculating bias for each user and each item 
training dataset  

**user bias**
```{r}
user_mean <- rowMeans(training, na.rm= TRUE)
user_bias <- user_mean-raw_average
user_bias
```

**item bias**
```{r}
item_mean <- colMeans(training, na.rm = TRUE)
item_bias <- item_mean- raw_average
item_bias
```

## Calculating baseline predictors 
every user-item combination. 

```{r}
baselineP <- function(n,m){
  3.5+ user_bias[n]+item_bias[m]
}
baseline_predictor <- data.frame(matrix(NA, ncol=6, nrow=6))
colnames(baseline_predictor)<- c("Marriott","Hilton","Best Western", "Hyatt", "Intercontinental","Wyndham")
rownames(baseline_predictor)<- c("user1","user2","user3","user4","user5","user6")
for (i in 1:6){
  for(j in 1:6){
    baseline_predictor[i,j] <- baselineP(i,j)
  }
}
baseline_predictor
```

## Calculating RMSE for the baseline predictors
Test dataset
```{r}
rmse_test_b <- sqrt(((test[1]-4.369565)^2 + (test[2]-4.369565)^2 +(test[3]-2.869565)^2 + (test[4]-3.536232)^2 +(test[5]-3.536232)^2 + (test[6]-3.569565)^2)/5)
rmse_test_b
```
Training dataset
```{r}
a<- (training -baseline_predictor)^2
a
mean_a <- sum(colSums(a, na.rm=TRUE, dims= 1))/23
rmse_training_b <- sqrt(mean_a)
rmse_training_b
```

## Summurizing results
Percentage of improvement of test dataset - 29.6%
```{r}
(1-(rmse_test_b/rmse_test))*100

```

Percentage of improvement of training dataset - 7.7%
```{r}
(1-(rmse_training_b/rmse_training))*100

```