---
title: "Data612 Project 2"
author: "Eunkyu Hahm"
date: "6/13/2020"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project 2 | Content-Based and Collaborative Filtering

### 1. load libraries
```{r}
library(tidyverse)
library(kableExtra)
library(Matrix)
library(knitr)
library(recommenderlab)
library(gridExtra)
```

### 2. Datasets
I use MovieLens small datasets: 100,000 ratings and 3,600 tag applications applied to 9,000 movies by 600 users.
```{r}
ratings_data <- read.csv("https://raw.githubusercontent.com/ekhahm/Data612/master/ratings.csv", stringsAsFactors = F)
movie_data <- read.csv("https://raw.githubusercontent.com/ekhahm/Data612/master/movies.csv", stringsAsFactors = F)
movie_data$movieId <- as.factor(movie_data$movieId)
head(ratings_data)
head(movie_data)
```


### 3. Subsetting data
```{r}
ratings_data$userId <- as.factor(ratings_data$userId)
movie_data$movieId <- as.factor(movie_data$movieId)
bmatrix <- as(ratings_data, "realRatingMatrix")
dim(bmatrix@data)
```

### 4. Similarity 

Users

```{r}
#Users
similarity_users <- similarity(bmatrix[1:20, ], method = "cosine", which = "users")
image(as.matrix(similarity_users), main = "User Similarity")
as.matrix(similarity_users)
```
Items

```{r}
#Items
similarity_items <- similarity(bmatrix[,1:20], method = "cosine", which = "items")
image(as.matrix(similarity_items), main = "Movie Similarity")
as.matrix(similarity_items)
```


### 5. Exploring the value of ratings

```{r}
#Distribution plot
ratings_dis <- as.data.frame(table(as.vector(bmatrix@data)))
ratings_dis
```


```{r}
# remove 0 ratings
ggplot(ratings_dis[-1,], aes(x = Var1, y= Freq)) + 
  geom_bar(stat= "identity") +
  geom_text(aes(label=Freq), vjust=1.6, color="white", size=3.5)+
  labs(title = "Distribution of the ratings", y = "Freq", x = "Ratings") +
  theme_minimal()
```
### 6. Exploring which movies have been viewed
```{r}
ratings_count <- colCounts(bmatrix)
viewed_movie <- data.frame(movieId = names(ratings_count), viewed = ratings_count)
top_movie <- inner_join(viewed_movie, movie_data, by = "movieId")%>%
  arrange(desc(viewed))%>%
  select(-movieId)%>%
  head(10)
top_movie
ggplot(top_movie, aes(x= reorder(title, -viewed), y=viewed))+ geom_bar(stat = "identity", fill = "#f68060")+ theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### 7. Exploring the average ratings

```{r}
average_ratings <- data.frame("average" = colMeans(bmatrix))
ggplot(average_ratings, aes(x=average))+ stat_bin(binwidth = 0.1) +
   ggtitle("Distribution of the average movie rating")

```

### 8. Visualizing the matrix
```{r}
image(bmatrix[1:100, 1:100], main = "Heatmap of the rating matrix")
```

### 9. Recommendation model
#### Model 1 (Item-Item Collaborative Filtering)
```{r}
#training and testing dataset
dim(bmatrix)
ratings_movies <- bmatrix[rowCounts(bmatrix) > 100, colCounts(bmatrix) > 100]
which_train <- sample(x = c(TRUE, FALSE), size = nrow(ratings_movies), replace = TRUE, prob = c(0.8, 0.2))
train <- ratings_movies[which_train, ]
test <- ratings_movies[!which_train, ]
```

train modelling
```{r}
Imodel <- Recommender(data=train, method="IBCF")
Imodel
```

Predicting 
```{r}
recc_predicted <- predict(Imodel, newdata= test, n=6)
recc_predicted@items[[1]]
recc_matrix1 <- sapply(recc_predicted@items, function(x) {colnames(ratings_movies)[x]})
recc_matrix1[, 1:3] %>% kable() %>% kable_styling(full_width = T) 
```

#### Model 2 (User-User Collaborative Filtering)
train modelling
```{r}
Umodel <- Recommender(data=train, method="UBCF")
Umodel
```

Predicting 
```{r}
recc_predicted <- predict(Umodel, newdata= test, n=6)
recc_predicted@items[[1]]
recc_matrix1 <- sapply(recc_predicted@items, function(x) {colnames(ratings_movies)[x]})
recc_matrix1[, 1:3] %>% kable() %>% kable_styling(full_width = T) 
```